Concurrent Programming with Harmony

### 2020-07-30 book/paper.tex
Fixed typo

Some programming languages favor a different way of implementing
synchronization using so-called \emph{actors}~\cite{HBS73}.  Actors are
processes that have only private memory and communicate through
\emph{message passing}.
See \autoref{fig:actorpic} for an illustration.
Given that there is no shared memory in the actor model (other than the message
queues, which have built-in synchronization), there is no need
for critical sections.  Instead, some sequential process owns a particular
piece of data and other processes access it by sending request messages
to the process and optionally waiting for response messages.  Each process
handles one message at a time, serializing all access to the data it owns.
As message queues are FCFS, starvation is prevented.

@@

\autoref{fig:actor} shows a Harmony solution to the
``multiple producer / multiple consumer problem'' (with an unbounded buffer)
using an actor (as illustrated in \autoref{fig:actorpic}.
\texttt{pc\_actor} is essentially a matchmaker: it matches
\texttt{.produce} requests with \texttt{.consume} requests.
Sometimes it may have buffered \texttt{.consume} requests and is waiting for
\texttt{.produce} requests, and sometimes it is vice versa.
The variable \textit{balance} specifies what the situation is: if it is positive
then it has buffered \texttt{.produce} requests;
if it is negative
then it has buffered \texttt{.consume} requests.
Note that both \textit{balance} and \textit{requests} are variables that are
local to the \texttt{pc\_actor} process.

@@

\section*{Exercises}
\begin{problems}
\item \label{ex:cltsvr} A popular model is the \emph{client/server} model.
Here a single actor implements some service, like computing the square of a
number.
A client can send a message containing an integer to the server, and the
server returns a response message containing the square of that message.
Implement this.
\item Actors and message queues are good for building pipelines.
Develop a pipeline that computes Mersenne primes (primes that are one less
than a factor of two).  Write four actors:
\begin{enumerate}
\item an actor that generates a sequence of integers 1 through \texttt{N};
\item an actor that receives integers and forwards only those that are prime;
\item an actor that receives integers and forwards only those that are one
less than a factor of two;
\item an actor that receives integers but otherwise ignores them.
\end{enumerate}
Configure two versions of the pipeline, one that first checks if a number
is prime and then if it is one less than a factor of two, the other
in the opposite order.  Which do you think is better?
\end{problems}

>>>>

Some programming languages favor a different way of implementing
synchronization using so-called \emph{actors}~\cite{HBS73}.  Actors are
processes that have only private memory and communicate through
\emph{message passing}.
See \autoref{fig:actorpic} for an illustration.
Given that there is no shared memory in the actor model (other than the message
queues, which have built-in synchronization), there is no need
for critical sections.  Instead, some sequential process owns a particular
piece of data and other processes access it by sending request messages
to the process and optionally waiting for response messages.  Each process
handles one message at a time, serializing all access to the data it owns.
As message queues are FCFS (First-Come-First-Served), starvation is prevented.

@@

\autoref{fig:actor} shows a Harmony solution to the
``multiple producer / multiple consumer problem'' (with an unbounded buffer)
using an actor (as illustrated in \autoref{fig:actorpic}).
\texttt{pc\_actor} is essentially a matchmaker: it matches
\texttt{.produce} requests with \texttt{.consume} requests.
Sometimes it may have buffered \texttt{.consume} requests and is waiting for
\texttt{.produce} requests, and sometimes it is vice versa.
The variable \textit{balance} specifies what the situation is: if it is positive
then it has buffered \texttt{.produce} requests;
if it is negative
then it has buffered \texttt{.consume} requests.
Note that both \textit{balance} and \textit{requests} are variables that are
local to the \texttt{pc\_actor} process.

@@

\section*{Exercises}
\begin{problems}
\item \label{ex:cltsvr} A popular model is the \emph{client/server} model.
Here a single actor implements some service, like computing the square of a
number.
A client can send a message containing an integer to the server, and the
server returns a response message containing the square of that integer.
Implement this.
\item Actors and message queues are good for building pipelines.
Develop a pipeline that computes Mersenne primes (primes that are one less
than a power of two).  Write four actors:
\begin{enumerate}
\item an actor that generates a sequence of integers 1 through \texttt{N};
\item an actor that receives integers and forwards only those that are prime;
\item an actor that receives integers and forwards only those that are one
less than a power of two;
\item an actor that receives integers but otherwise ignores them.
\end{enumerate}
Configure two versions of the pipeline, one that first checks if a number
is prime and then if it is one less than a power of two, the other
in the opposite order.  Which do you think is better?
\end{problems}
